### 2016-02-12 text/Chapter04.tex
Normalize data-flow and dataflow

In the previous chapter, we looked at three important examples of systems that track whole-context 
properties. The type systems for whole-context liveness (Section~\ref{sec:applications-flat-live}) 
and whole-context data-flow (Section~\ref{sec:applications-flat-dataflow}) have a very similar 
structure. First, their lambda abstraction duplicates the requirements. Given a body with context 
requirements $\cclrd{r}$, the declaration site context \emph{as well as} the function arrow are
annotated with $\cclrd{r}$. Second, their application arises from the combination of \emph{sequential} 
and \emph{pointwise} composition.

@@

Recall the lambda abstraction rules for the implicit parameters system (annotating the context
with sets of required parameters) and the data-flow system (annotating the context with the
number of past required values):
%
\begin{equation*}
\tyrule{param}
  {\coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r} \cup \cclrd{s}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 }
\;
\tyrule{df1}
  {\coctx{\Gamma, x\!:\!\tau_1}{\cclrd{n}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{n}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{n}} \tau_2 }
\end{equation*}
%
In order to capture both systems using a single calculus, we need a way of unifying the two
systems. For the data-flow system, this can be achieved by over-approximating the number of 
required past elements:
%
\begin{equation*}
\tyrule{df2}
  {\coctx{\Gamma, x\!:\!\tau_1}{\cclrd{\textnormal{min}(n, m)}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{n}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{m}} \tau_2 }
\end{equation*}
%
The rule (\emph{df1}) is admissible in a system that includes the (\emph{df2}) rule. 
Furthermore, if we include sub-typing rule (on annotations of functions) and sub-coeffecting rule (on 
annotations of contexts), then the reverse is also true -- because 
$\cclrd{\textit{min}(n, m)} \leq \cclrd{m}$ and $\cclrd{\textit{min}(n, m)} \leq \cclrd{n}$.
In other words (\emph{df1}) is more precise, but (\emph{df2}) gives a sound over-approximation
with a structure that can be unified with (\emph{param}).

@@

\noindent
In two of the three systems, some of the operators of the flat coeffect algebra coincide, but
in the data-flow system all three are distinct. Similarly, the two special elements  
coincide in some, but not all systems. The required laws are motivated by the aim to capture
common properties of the three examples, without unnecessarily restricting the system:

@@

\noindent
It is worth noting that, in some of the systems, the operators $\cpar$ and $\czip$ are the least 
upper bound and the greatest lower bounds of a lattice. For example, in data-flow computations, they 
are \emph{max} and \emph{min} respectively. However, this duality does not hold for implicit parameters
(we discuss lattice-based formulation of coeffects in Section~\ref{sec:unified-impl-semilattice}.

@@

The flat coeffect calculus generalizes the flat systems discussed in 
Section~\ref{sec:applications-flat} of the previous chapter. We can instantiate it to a specific
use just by providing a flat coeffect algebra. The following summary defines the systems for implicit 
parameters, liveness and data-flow. For the latter two, the general calculus has a lambda abstraction
that is compatible with those discussed in Chapter~\ref{ch:flat}, but includes implicit sub-coeffecting.

@@

\begin{example}[Data-flow]
In data-flow, context is annotated with natural numbers and the flat coeffect algebra is formed 
by $(\mathbb{N}, +, \mathit{max}, \mathit{min}, 0, 0, \leq)$.
\end{example}

\noindent
As discussed earlier, sequential composition $\cseq$ is represented by $+$ and pointwise 
composition $\cpar$ uses $\emph{max}$. For data-flow, we need a third separate operator for
lambda abstraction. Annotating the body with $\emph{min}(\cclrd{r}, \cclrd{s})$ ensures that
both call site and declaration site annotations are equal or greater than the annotation 
of the body. As with liveness, this allows over-approximation.

As required by the laws, $(\mathbb{N}, +, 0)$ and $(\mathbb{N}, \mathit{max}, 0)$ form monoids
and $(\mathbb{N}, \mathit{min})$ forms a band. Note that data-flow is our first example where 
$\cseq$ is not idempotent. The distributivity laws require the following to be the case:
$\mathit{max}(r,s) + t = \mathit{max}(r+t, s+t)$, which is easy to see.

A simple dataflow langauge includes an additional construct $\kvd{prev}$ for accessing the 
previous value in a stream with an additional typing rule that look as follows:
%
\begin{equation*}
\quad~ e ~::=~ \ldots \sep \kvd{next}~e
\end{equation*}
\begin{equation*}
\tyrule{prev}
  { \coctx{\Gamma}{\cclrd{n}} \vdash e : \tau }
  { \coctx{\Gamma}{\cclrd{n+1}} \vdash \kvd{prev}~e : \tau }
\end{equation*}
%
As a further example that was not covered earlier, it is also possible to combine liveness analysis
and data-flow. In the above calculus, $0$ denotes that we require the current value, but no previous
values. However, for constants, we do not even need the current value.

\begin{example}[Optimized data-flow]
In optimized data-flow, context is annotated with natural numbers extended with the $\bot$ element,
that is $\mathbb{N}_{\bot} = \{\bot, 0, 1, 2, 3, \ldots \}$ such that $\forall n \in \mathbb{N}. \bot \leq n$.
The flat coeffect algebra is $(\mathbb{N}_{\bot}, +, \mathit{max}, \mathit{min}, 0, \bot, \leq)$
where $m + n$ is $\bot$ whenever $m=\bot$ or $n=\bot$ and \emph{min}, \emph{max} treat $\bot$ as the
least element.
\end{example}

@@

In this section, we look at such shared syntactic properties. This accompanies the previous section, 
which provided a \emph{semantic} justification for the axioms of coeffect algebra with a 
\emph{syntactic} justification. Operationally, this section can also be viewed as providing a 
pathway to an operational semantics for two of our systems (implicit parameters and liveness),
which can be based on syntactic substitution. As we discuss later, the notion of context for
data-flow is more complex.

@@

The property holds for the three systems we consider -- for implicit parameters, this is 
an equality; for liveness and data-flow (which both use a lattice), the greatest lower bound 
is smaller than the least upper bound.

@@

\paragraph{Data-flow.} 
The type preservation property does not hold for data-flow. This case is similar to the example
shown earlier with effectful computations. As a minimal example, consider the substitution of 
a context-dependent expression $\kvd{prev}~z$ for a variable $y$ in a function $\lambda x.y$:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1,z\!:\!\tau_1}{0} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{0} \tau_2 & & (\textnormal{before}) \\
 \coctx{z\!:\!\tau_1}{1} &\narrow{\vdash}& \lambda x.\kvd{prev}~z : \tau_1 \xrightarrow{1} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
After the substitution, the coeffect of the body is $1$. The rule for lambda abstraction requires
that $1=\mathit{min}(r,s)$ and so the least solution is to set both $r,s$ to $1$. The substitution
this affects the coeffects attached both to the function type and the overall context.

@@

\paragraph{Implicit parameters.} In data-flow, there is no typing for the resulting expression that
preserves the type of the function. However, this is not the case for all systems. Consider substituting
an implicit parameter access $\ident{?p}$ for a free variable $y$ under a lambda:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1}{\emptyset} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{before}) \\
 \coctx{\emptyset}{ \{\ident{?p}\} } &\narrow{\vdash}& \lambda x.\ident{?p} : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
The above shows one possible typing of the body -- one that does not change the coeffects of the 
function type and attaches all additional coeffects (implicit parameters) to the context. In case
of implicit parameters (and, more generally, systems with set-like annotations) this is always 
possible.

@@

\noindent
Liveness is an example of top-pointed coeffects as variables are annotated with 
$\ident{L}$ and $\ident{D} \leq \ident{L}$, while implicit parameters and data-flow are examples
of bottom-pointed coeffects. For top-pointed flat coeffects, the substitution lemma holds without 
additional requirements:

@@

\paragraph{Expansion theorem.}
The $\eta$-expansion (local completeness) is similar to $\beta$-reduction (local soundness) in that
it holds for some flat coeffect systems, but not for all. Out of the examples we discuss, it 
holds for implicit parameters, but does not hold for liveness and data-flow.

Recall that $\eta$-expansion turns $e$ into $\lambda x.e~x$. In the case of liveness, the 
expression $e$ may require no variables (both immediate and latent coeffects are marked as
$\ident{D}$). However, the resulting expression $\lambda x.e~x$ accesses a variable, 
marking the context and function argument as live. In case of data-flow, the immediate coeffects 
are made larger by the lambda abstraction -- the context requirements of the function value are 
imposed on the declaration site of the new lambda abstraction. We remedy this limitation in 
the next chapter.

@@

\begin{table}
\begin{center}
\begin{tabular}{ | l | c | c |}
\hline
& \textbf{\footnotesize Definition\hspace{1em}} & \textbf{\footnotesize Simplified\hspace{1em}} \\ \hline
\hspace{-1em}{\footnotesize Implicit parameters} & $\cclrd{s} \cup (\cclrd{s} \cup \cclrd{r})$ & $\cclrd{s} \cup \cclrd{r}$ \\ \hline
\hspace{-1em}{\footnotesize Liveness} & $\cclrd{s} \sqcap (\cclrd{s} \sqcup \cclrd{r})$ & $\cclrd{s}$ \\ \hline
\hspace{-1em}{\footnotesize Data-flow} & $\mathit{max}(\cclrd{s}, \cclrd{s} + \cclrd{r})$ & $\cclrd{s} + \cclrd{r}$ \\ \hline
\end{tabular}
\end{center}

@@

To see that the simplified annotations are more precise, assume that we used arbitrary 
splitting $\cclrd{s} = \cclrd{s_1}\,\czip\,\cclrd{s_2}$ rather than idempotence. The
``Definition'' column would use $\cclrd{s_1}$ and $\cclrd{s_2}$ for the first and second 
$\cclrd{s}$, respectively. The corresponding simplified annotation would have 
$\cclrd{s_1}\,\czip\,\cclrd{s_2}$ instead of $\cclrd{s}$. For all our systems, the 
simplified annotation (on the right) is more precise than the original (on the left):
%
\begin{equation*}
\begin{array}{rclll}
\cclrd{s_1} \cup (\cclrd{s_2} \cup \cclrd{r}) &\supseteq& (\cclrd{s_1} \cup \cclrd{s_2}) \cup \cclrd{r} 
  && \textnormal{(implicit parameters)}\\
\cclrd{s_1} \sqcap (\cclrd{s_2} \sqcup \cclrd{r}) &\sqsupseteq&  (\cclrd{s_1} \sqcap \cclrd{s_2}) 
  && \textnormal{(liveness)} \\
\mathit{max}(\cclrd{s_1}, \cclrd{s_2} + \cclrd{r}) &\geq& \mathit{min}(\cclrd{s_1}, \cclrd{s_2}) + \cclrd{r} 
  &\quad& \textnormal{(data-flow)} \\
\end{array}
\end{equation*}
%
In other words, the inequality states that using idempotence, we get a more precise typing.
Using the $\cgeq$ operator this property can be expressed using the abstract operators of the flat coeffect algebra as:
%
\begin{equation*}
\cclrd{s_1} \;\cpar\; (\cclrd{s_2} \,\cseq\, \cclrd{r}) \;\cgeq\; 
  (\cclrd{s_1} \,\czip\, \cclrd{s_2}) \;\cpar\; ((\cclrd{s_1} \,\czip\, \cclrd{s_2}) \,\cseq\, \cclrd{r})
\end{equation*}
%
This property cannot be proved from other properties of the flat coeffect algebra. To make
the flat coeffect system as general as possible, we do not \emph{in general} require it as
an additional axiom, although the above examples provide reasonable basis for requiring 
that the specialized annotation for let binding is the least possible annotation for the 
expression $(\lambda x.e_2)~e_1$.

@@

In Section~\ref{sec:flat-calculus-lambda}, we discussed how to reconcile two typings for
lambda abstraction -- for implicit parameters, the lambda function needs to split context 
requirements using $\cclrd{r} \cup \cclrd{s}$, but for data-flow and liveness it suffices to 
duplicate the requirement $\cclrd{r}$ of the body. We introduced the $\czip$ operation as 
a way of providing the additional abstraction.

@@

\noindent
The practical consequence of the Remark~\ref{thm:flat-alt-abs} is that, for strictly
oriented coeffect calculi (such as our liveness and data-flow computations), we can use
the (\emph{idabs}) rule and get an equivalent type system. This alternative formulation
removes the non-determinism of type checking that arises from the splitting of context 
requirements in the original (\emph{abs}) rule.

@@

The flat coeffect calculus is parameterized by a \emph{flat coeffect algebra} that captures
the structure of the information tracked by the type system. We instantiated the system to 
capture three specific systems, namely liveness, data-flow and implicit parameters. However,
the system is more general and an capture numerous other applications outlined in 
Section~\ref{sec:applications-flat}.

>>>>

In the previous chapter, we looked at three important examples of systems that track whole-context 
properties. The type systems for whole-context liveness (Section~\ref{sec:applications-flat-live}) 
and whole-context dataflow (Section~\ref{sec:applications-flat-dataflow}) have a very similar 
structure. First, their lambda abstraction duplicates the requirements. Given a body with context 
requirements $\cclrd{r}$, the declaration site context \emph{as well as} the function arrow are
annotated with $\cclrd{r}$. Second, their application arises from the combination of \emph{sequential} 
and \emph{pointwise} composition.

@@

Recall the lambda abstraction rules for the implicit parameters system (annotating the context
with sets of required parameters) and the dataflow system (annotating the context with the
number of past required values):
%
\begin{equation*}
\tyrule{param}
  {\coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r} \cup \cclrd{s}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 }
\;
\tyrule{df1}
  {\coctx{\Gamma, x\!:\!\tau_1}{\cclrd{n}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{n}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{n}} \tau_2 }
\end{equation*}
%
In order to capture both systems using a single calculus, we need a way of unifying the two
systems. For the dataflow system, this can be achieved by over-approximating the number of 
required past elements:
%
\begin{equation*}
\tyrule{df2}
  {\coctx{\Gamma, x\!:\!\tau_1}{\cclrd{\textnormal{min}(n, m)}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{n}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{m}} \tau_2 }
\end{equation*}
%
The rule (\emph{df1}) is admissible in a system that includes the (\emph{df2}) rule. 
Furthermore, if we include sub-typing rule (on annotations of functions) and sub-coeffecting rule (on 
annotations of contexts), then the reverse is also true -- because 
$\cclrd{\textit{min}(n, m)} \leq \cclrd{m}$ and $\cclrd{\textit{min}(n, m)} \leq \cclrd{n}$.
In other words (\emph{df1}) is more precise, but (\emph{df2}) gives a sound over-approximation
with a structure that can be unified with (\emph{param}).

@@

\noindent
In two of the three systems, some of the operators of the flat coeffect algebra coincide, but
in the dataflow system all three are distinct. Similarly, the two special elements  
coincide in some, but not all systems. The required laws are motivated by the aim to capture
common properties of the three examples, without unnecessarily restricting the system:

@@

\noindent
It is worth noting that, in some of the systems, the operators $\cpar$ and $\czip$ are the least 
upper bound and the greatest lower bounds of a lattice. For example, in dataflow computations, they 
are \emph{max} and \emph{min} respectively. However, this duality does not hold for implicit parameters
(we discuss lattice-based formulation of coeffects in Section~\ref{sec:unified-impl-semilattice}.

@@

The flat coeffect calculus generalizes the flat systems discussed in 
Section~\ref{sec:applications-flat} of the previous chapter. We can instantiate it to a specific
use just by providing a flat coeffect algebra. The following summary defines the systems for implicit 
parameters, liveness and dataflow. For the latter two, the general calculus has a lambda abstraction
that is compatible with those discussed in Chapter~\ref{ch:flat}, but includes implicit sub-coeffecting.

@@

\begin{example}[Dataflow]
In dataflow, context is annotated with natural numbers and the flat coeffect algebra is formed 
by $(\mathbb{N}, +, \mathit{max}, \mathit{min}, 0, 0, \leq)$.
\end{example}

\noindent
As discussed earlier, sequential composition $\cseq$ is represented by $+$ and pointwise 
composition $\cpar$ uses $\emph{max}$. For dataflow, we need a third separate operator for
lambda abstraction. Annotating the body with $\emph{min}(\cclrd{r}, \cclrd{s})$ ensures that
both call site and declaration site annotations are equal or greater than the annotation 
of the body. As with liveness, this allows over-approximation.

As required by the laws, $(\mathbb{N}, +, 0)$ and $(\mathbb{N}, \mathit{max}, 0)$ form monoids
and $(\mathbb{N}, \mathit{min})$ forms a band. Note that dataflow is our first example where 
$\cseq$ is not idempotent. The distributivity laws require the following to be the case:
$\mathit{max}(r,s) + t = \mathit{max}(r+t, s+t)$, which is easy to see.

A simple dataflow langauge includes an additional construct $\kvd{prev}$ for accessing the 
previous value in a stream with an additional typing rule that look as follows:
%
\begin{equation*}
\quad~ e ~::=~ \ldots \sep \kvd{next}~e
\end{equation*}
\begin{equation*}
\tyrule{prev}
  { \coctx{\Gamma}{\cclrd{n}} \vdash e : \tau }
  { \coctx{\Gamma}{\cclrd{n+1}} \vdash \kvd{prev}~e : \tau }
\end{equation*}
%
As a further example that was not covered earlier, it is also possible to combine liveness analysis
and dataflow. In the above calculus, $0$ denotes that we require the current value, but no previous
values. However, for constants, we do not even need the current value.

\begin{example}[Optimized dataflow]
In optimized dataflow, context is annotated with natural numbers extended with the $\bot$ element,
that is $\mathbb{N}_{\bot} = \{\bot, 0, 1, 2, 3, \ldots \}$ such that $\forall n \in \mathbb{N}. \bot \leq n$.
The flat coeffect algebra is $(\mathbb{N}_{\bot}, +, \mathit{max}, \mathit{min}, 0, \bot, \leq)$
where $m + n$ is $\bot$ whenever $m=\bot$ or $n=\bot$ and \emph{min}, \emph{max} treat $\bot$ as the
least element.
\end{example}

@@

In this section, we look at such shared syntactic properties. This accompanies the previous section, 
which provided a \emph{semantic} justification for the axioms of coeffect algebra with a 
\emph{syntactic} justification. Operationally, this section can also be viewed as providing a 
pathway to an operational semantics for two of our systems (implicit parameters and liveness),
which can be based on syntactic substitution. As we discuss later, the notion of context for
dataflow is more complex.

@@

The property holds for the three systems we consider -- for implicit parameters, this is 
an equality; for liveness and dataflow (which both use a lattice), the greatest lower bound 
is smaller than the least upper bound.

@@

\paragraph{Dataflow.} 
The type preservation property does not hold for dataflow. This case is similar to the example
shown earlier with effectful computations. As a minimal example, consider the substitution of 
a context-dependent expression $\kvd{prev}~z$ for a variable $y$ in a function $\lambda x.y$:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1,z\!:\!\tau_1}{0} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{0} \tau_2 & & (\textnormal{before}) \\
 \coctx{z\!:\!\tau_1}{1} &\narrow{\vdash}& \lambda x.\kvd{prev}~z : \tau_1 \xrightarrow{1} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
After the substitution, the coeffect of the body is $1$. The rule for lambda abstraction requires
that $1=\mathit{min}(r,s)$ and so the least solution is to set both $r,s$ to $1$. The substitution
this affects the coeffects attached both to the function type and the overall context.

@@

\paragraph{Implicit parameters.} In dataflow, there is no typing for the resulting expression that
preserves the type of the function. However, this is not the case for all systems. Consider substituting
an implicit parameter access $\ident{?p}$ for a free variable $y$ under a lambda:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1}{\emptyset} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{before}) \\
 \coctx{\emptyset}{ \{\ident{?p}\} } &\narrow{\vdash}& \lambda x.\ident{?p} : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
The above shows one possible typing of the body -- one that does not change the coeffects of the 
function type and attaches all additional coeffects (implicit parameters) to the context. In case
of implicit parameters (and, more generally, systems with set-like annotations) this is always 
possible.

@@

\noindent
Liveness is an example of top-pointed coeffects as variables are annotated with 
$\ident{L}$ and $\ident{D} \leq \ident{L}$, while implicit parameters and dataflow are examples
of bottom-pointed coeffects. For top-pointed flat coeffects, the substitution lemma holds without 
additional requirements:

@@

\paragraph{Expansion theorem.}
The $\eta$-expansion (local completeness) is similar to $\beta$-reduction (local soundness) in that
it holds for some flat coeffect systems, but not for all. Out of the examples we discuss, it 
holds for implicit parameters, but does not hold for liveness and dataflow.

Recall that $\eta$-expansion turns $e$ into $\lambda x.e~x$. In the case of liveness, the 
expression $e$ may require no variables (both immediate and latent coeffects are marked as
$\ident{D}$). However, the resulting expression $\lambda x.e~x$ accesses a variable, 
marking the context and function argument as live. In case of dataflow, the immediate coeffects 
are made larger by the lambda abstraction -- the context requirements of the function value are 
imposed on the declaration site of the new lambda abstraction. We remedy this limitation in 
the next chapter.

@@

\begin{table}
\begin{center}
\begin{tabular}{ | l | c | c |}
\hline
& \textbf{\footnotesize Definition\hspace{1em}} & \textbf{\footnotesize Simplified\hspace{1em}} \\ \hline
\hspace{-1em}{\footnotesize Implicit parameters} & $\cclrd{s} \cup (\cclrd{s} \cup \cclrd{r})$ & $\cclrd{s} \cup \cclrd{r}$ \\ \hline
\hspace{-1em}{\footnotesize Liveness} & $\cclrd{s} \sqcap (\cclrd{s} \sqcup \cclrd{r})$ & $\cclrd{s}$ \\ \hline
\hspace{-1em}{\footnotesize Dataflow} & $\mathit{max}(\cclrd{s}, \cclrd{s} + \cclrd{r})$ & $\cclrd{s} + \cclrd{r}$ \\ \hline
\end{tabular}
\end{center}

@@

To see that the simplified annotations are more precise, assume that we used arbitrary 
splitting $\cclrd{s} = \cclrd{s_1}\,\czip\,\cclrd{s_2}$ rather than idempotence. The
``Definition'' column would use $\cclrd{s_1}$ and $\cclrd{s_2}$ for the first and second 
$\cclrd{s}$, respectively. The corresponding simplified annotation would have 
$\cclrd{s_1}\,\czip\,\cclrd{s_2}$ instead of $\cclrd{s}$. For all our systems, the 
simplified annotation (on the right) is more precise than the original (on the left):
%
\begin{equation*}
\begin{array}{rclll}
\cclrd{s_1} \cup (\cclrd{s_2} \cup \cclrd{r}) &\supseteq& (\cclrd{s_1} \cup \cclrd{s_2}) \cup \cclrd{r} 
  && \textnormal{(implicit parameters)}\\
\cclrd{s_1} \sqcap (\cclrd{s_2} \sqcup \cclrd{r}) &\sqsupseteq&  (\cclrd{s_1} \sqcap \cclrd{s_2}) 
  && \textnormal{(liveness)} \\
\mathit{max}(\cclrd{s_1}, \cclrd{s_2} + \cclrd{r}) &\geq& \mathit{min}(\cclrd{s_1}, \cclrd{s_2}) + \cclrd{r} 
  &\quad& \textnormal{(dataflow)} \\
\end{array}
\end{equation*}
%
In other words, the inequality states that using idempotence, we get a more precise typing.
Using the $\cgeq$ operator this property can be expressed using the abstract operators of the flat coeffect algebra as:
%
\begin{equation*}
\cclrd{s_1} \;\cpar\; (\cclrd{s_2} \,\cseq\, \cclrd{r}) \;\cgeq\; 
  (\cclrd{s_1} \,\czip\, \cclrd{s_2}) \;\cpar\; ((\cclrd{s_1} \,\czip\, \cclrd{s_2}) \,\cseq\, \cclrd{r})
\end{equation*}
%
This property cannot be proved from other properties of the flat coeffect algebra. To make
the flat coeffect system as general as possible, we do not \emph{in general} require it as
an additional axiom, although the above examples provide reasonable basis for requiring 
that the specialized annotation for let binding is the least possible annotation for the 
expression $(\lambda x.e_2)~e_1$.

@@

In Section~\ref{sec:flat-calculus-lambda}, we discussed how to reconcile two typings for
lambda abstraction -- for implicit parameters, the lambda function needs to split context 
requirements using $\cclrd{r} \cup \cclrd{s}$, but for dataflow and liveness it suffices to 
duplicate the requirement $\cclrd{r}$ of the body. We introduced the $\czip$ operation as 
a way of providing the additional abstraction.

@@

\noindent
The practical consequence of the Remark~\ref{thm:flat-alt-abs} is that, for strictly
oriented coeffect calculi (such as our liveness and dataflow computations), we can use
the (\emph{idabs}) rule and get an equivalent type system. This alternative formulation
removes the non-determinism of type checking that arises from the splitting of context 
requirements in the original (\emph{abs}) rule.

@@

The flat coeffect calculus is parameterized by a \emph{flat coeffect algebra} that captures
the structure of the information tracked by the type system. We instantiated the system to 
capture three specific systems, namely liveness, dataflow and implicit parameters. However,
the system is more general and an capture numerous other applications outlined in 
Section~\ref{sec:applications-flat}.
